---
title: "ARIMA results"
author: "Theresia Deiss"
date: "`r Sys.Date()`"
output: html_document
---

## Loading packages

```{r}
library("latex2exp")
library("grDevices")
library("tseries")
library("forecast")

setwd("C:\\Users\\deiss\\Desktop\\Semester\\SDSU_Spring_Semester\\Y_Master_Thesis\\Coding\\Plots")
```


## Trend analysis

```{r}

pc1 = data.pca$x[,1]
pc1.dif = diff(pc1)

# Make TS
pc1 <- ts(pc1,  frequency=252, start = 2015)
pc1.dif <- ts(pc1.dif,  frequency=252, start = 2015)

# Plot pc1 and 1st difference
jpeg(file = "Chap4_PC1.jpeg", width = 7, height = 8, units = 'in', res = 300)
par(mfrow=c(2,1))
plot(pc1, xlab = "Year", ylab = TeX("$PC_1$"), col="blue", main = TeX("\\textbf{PC$_{1}$ over Time}"))
#title( TeX("\\textbf{S&P 500 price ($Z_t$)}"))
plot(pc1.dif, xlab = "Year", ylab = TeX("diff($PC_1$)"), col="blue", main = TeX("\\textbf{First Difference of PC$_{1}$ over Time}"))
#title( TeX("\\textbf{S&P 500 return ($\\textbf{1}^{st}$ difference $Z_t-Z_{t-1}$)}"))
dev.off()

```


## Unit root analysis: test stationarity

Augmented Dickey-Fuller Test:
H0: The time series is non-stationary (equiv: TS has a unit root)
HA: The time series is stationary.
If p-value<0.05: reject H0, hence TS is stationary.

```{r}

# Augmented Dickey-Fuller Test
adf.test(pc1)
adf.test(pc1.dif)

```
# Analysis of optimal parameters

```{r}
# Analysis of ACF and PACF

pc1.dif_scale <- ts(pc1.dif,  frequency=1)

jpeg(file = "Chap4_ACF_PC.jpeg", width = 7, height = 6.5, units = 'in', res = 300)
par(mfrow=c(2,1))
#acf(pc1.dif,  main = TeX("\\textbf{ACF of $1^{st}$ Difference}"))
acf(pc1.dif_scale,  main = TeX("\\textbf{ACF of First Difference}"), lag.max = 20)
acf(pc1.dif_scale, type="partial", main = TeX("\\textbf{PACF of First Difference}"), lag.max = 20)
dev.off()

acf(pc1.dif)
acf(pc1.dif, type="partial")

acf(pc1.dif^2)
```

# Model fit

```{r}

# Model ARIMA211 -- scaled
pc1_scale <- ts(pc1, frequency = 1)
fit211_scale <- arima(pc1_scale, order=c(2,1,1))
jpeg(file = "Chap4_ARIMA211.jpeg", width = 7, height = 6, units = 'in', res = 300)
tsdiag(fit211_scale, gof.lag = 10)
dev.off()

# Model ARIMA211 -- not scaled
fit211 <- arima(pc1, order=c(2,1,1))
fit211
autoplot(fit211)

res = fit211$residuals
acf(res)
acf((res-mean(res))^2)

# Fit of best guess
fit010 <- arima(pc1_scale, order=c(0,1,0))
jpeg(file = "Chap4_ARIMA010.jpeg", width = 7, height = 6, units = 'in', res = 300)
tsdiag(fit010)
dev.off()

fit010
autoplot(fit010)

# Best guess of auto.arima
auto.arima(pc1, ic = c("aicc", "aic", "bic"))

```

## In-sample performance

```{r}
res = fit211$residuals
pred_pc1 = pc1-res

jpeg(file = "Chap4_ARIMA211_InSample.jpeg", width = 7, height = 4.5, units = 'in', res = 300)
plot(pc1, xlab = "Year", ylab = TeX("$PC_1$"), main = "ARIMA(2,1,1):\nIn-Sample Performance", col="blue", lty=1, lwd = 2)
lines(pred_pc1, col="red",lty=1)
lines(res, col="grey",lty=1)
legend("bottomright", legend = c( TeX("Actual $PC_1$"), TeX("Predicted $PC_1$"),"Residuals"), col = c("blue", "red", "grey"), lty = c(1,1,1), bty="n")
dev.off()

# In-sample performance values
sqrt(mean((fit211$residuals)^2))               # RMSE
mean(abs((fit211$residuals)/pc1)) * 100        # MAPE
1/length(pc1)*sum(abs(fit211$residuals))       # MAE

# Jarque-Bera statistic: check if residuals are normal
jarque.bera.test(fit211$residuals)          # p-value small, res not normal! bad
# compare this with qq-plot: see residuals are not normal at the ends..
qqnorm(fit211$residuals)

```

## Out of sample performance: Forecast for one sample period

```{r}

# Number of days to forecast:
testPeriod.length <- 12

# Index of end date for sampling period
pc1.length   <- length(pc1)
pc1.time     <- 2015+c(1:pc1.length)/252
startMonth   <- 1
endMonth     <- 1334    # corresponds to March 12th 2020
samplePeriod <- ts(pc1[startMonth:endMonth], start=2015, frequency=252)
samplePeriod.length <- endMonth-startMonth+1

# Fit ARIMA(211)
fit.sample <- arima(samplePeriod, order=c(2,1,1))
fit.sample

# Forecasts for testing period 
pred <- predict(fit.sample, n.ahead = testPeriod.length)

# Out of sample test ARIMA211 for PC1
forecasts <- pred$pred
observations <- pc1[(endMonth+1):(endMonth+testPeriod.length)]
diff <- forecasts-observations

# GRAPH: SECTION OF SAMPLING PERIOD + FORECASTS

# Decide section to plot:
# startDay.sec <- pc1.time[1305]
# startDay.index <- which(pc1.time == startDay.sec)
startDay.index <- 1284 # corresponds to Jan 2020 
sec.length <- length(pc1[startDay.index:(endMonth+testPeriod.length)])
sec.time <- pc1.time[startDay.index:(endMonth+testPeriod.length)]
samplePeriod.sec <- samplePeriod[(startDay.index):samplePeriod.length]


jpeg(file = "Chap4_ARIMA211_OutOfSample.jpeg", width = 7, height = 4.5, units = 'in', res = 300)
par(mfrow=c(1,1))
plot(sec.time,c(samplePeriod.sec,pred$pred), ylim=c(0,5), lty=1, type="n", xaxt = "n", xlab = "Time", ylab = TeX("$PC_1$"))
lines(sec.time[1:(sec.length-testPeriod.length)],samplePeriod.sec, lty=1)
points(sec.time[(sec.length-testPeriod.length+1):sec.length], pred$pred, pch=20, col = "red") 
lines(sec.time[(sec.length-testPeriod.length+1):sec.length], pred$pred + 1.96*pred$se, lty=2, col = "blue") 
lines(sec.time[(sec.length-testPeriod.length+1):sec.length], pred$pred - 1.96*pred$se, lty=2, col = "blue") 
points(sec.time[(sec.length-testPeriod.length+1):sec.length], pc1[(endMonth+1):(endMonth+testPeriod.length)], pch=20) 
axis(1, at = c(pc1.time[1284],pc1.time[1305], pc1.time[1334]),  labels = c("Jan 2020", "Feb 2020", "March 12th 2020"))
#axis(1, at = c(pc1.time[1284],pc1.time[1305], pc1.time[1325], pc1.time[1334]),  labels = c("Jan 2020", "Feb 2020","March 20", "March 12th 20"))
title("ARIMA(2,1,1):\nOut-of-Sample Performance")
legend("bottomright", legend = c( "Observation", "Forecast", "95% CI", TeX("$PC_1$")), col = c("black", "red", "blue"),lty=c(NA,NA,2,1),  pch = c(20,20, NA, NA), bty="n" ) #, horiz=TRUE)
dev.off()

```
## Some data for ARIMA010

```{r}

res010 = fit010$residuals
pred_pc1_ARIMA010 = pc1-res010

fit.sample_ARIMA010 <- arima(samplePeriod, order=c(0,1,0))
pred_ARIMA010 <- predict(fit.sample_ARIMA010, n.ahead = testPeriod.length)

```



